---
title: "Homework 7"
author: "Yuvraj Jain"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  1,4 (Pg. 358); 1,4 (Pgs. 371)

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
library(modelr)
options(na.action = na.warn)
```


# Chapter 18 (Pg. 358)

##  Exercise 1:

Instead of using `lm()` to to fit a straight line, you can `loess()` to fit a smooth curve. Repeat the process of model fitting, grid generation, predictions, and visualization on `sim1` using `loess()` instead of `lm()`. How does the result compare to `geom_smooth()`?

### Model Fitting:

```{r}
sim1_loess <- loess(y ~ x, data = sim1)
sim1_loess
```

### Generating Grid:

```{r}
grid <- sim1 %>%
  data_grid(x)
grid
```

### Adding Predictions:

```{r}
grid <- grid %>%
  add_predictions(sim1_loess)
grid
```

### Visualization:

```{r}
ggplot() +
  geom_point(sim1, mapping = aes(x = x, y = y)) + 
  geom_line(
    data = grid,
    mapping = aes(x = x, y = pred),
    color = "red",
    linewidth = 1
  )
```

The predictions of loess are the same as the default method for geom_smooth() because geom_smooth() uses loess() by default; the message even tells us that.

```{r}
ggplot() + 
  geom_point(sim1, mapping = aes(x = x, y = y)) +
  geom_smooth(grid, mapping = aes(x = x, y = pred))
```

##  Exercise 4:

Why might you want to look at a frequency polygon of absolute residuals? What are the pros and cons compared to looking at the raw residuals?

### Pros of looking at a frequency polygon of absolute residuals:

* Provides a clearer picture of the overall accuracy of the model.
* Captures the magnitude of errors, allowing us to see if the model is consistently making predictions that are off by a certain amount.
* Helps to identify patterns or trends in the errors, such as over- or under-prediction for certain values.

### Cons of looking at a frequency polygon of absolute residuals:

* Does not provide information about the direction of errors, which may be important in certain contexts.
* Does not allow for the identification of systematic biases, such as a tendency to consistently over- or under-predict values in a certain range.
* May be less intuitive or difficult to interpret.

# Chapter 18 (Pg. 371)

##  Exercise 1:

What happens if you repeat the analysis of sim2 using a model without an intercept? What happens to the model equation? What happens to the predictions?

Removing the intercept from our model:

```{r}
mod2a <- lm(y ~ x - 1, data = sim2)
mod2 <- lm(y ~ x, data = sim2)
```

The predictions are exactly the same in the models with and without an intercept:

```{r}
grid <- sim2 %>%
  data_grid(x) %>%
  spread_predictions(mod2, mod2a)
grid
```

##  Exercise 4: 

For sim4, which of mod1 and mod2 is better? I think mod2 does a slightly better job at removing patterns, but it’s pretty subtle. Can you come up with a plot to support my claim?

Based on the models `mod1` and `mod2` written in the readings before the exercise, we know that:

```{r}
mod1 <- lm(y ~ x1 + x2, data = sim4)
mod2 <- lm(y ~ x1 * x2, data = sim4)
sim4_mods <- gather_residuals(sim4, mod1, mod2)
```

I can make a frequency polygon plot for the residuals of mod1 and mod2 to make a more appropriate comparison between the two.
Here is the frequency polygon below:

```{r}
ggplot(sim4_mods, aes(x = abs(resid), color = model)) +
  geom_freqpoly(binwidth = 0.5) +
  geom_rug()
```

From the above plot, we can see that mod2 has slightly lesser residual values than mod1, and this confirms that `mod2` does a slightly better job at removing patterns.
