---
title: "Ch1"
output: html_document
date: "2023-01-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

To access the datasets, help pages, and functions that we will use in this chapter, load the tidyverse by running this code:

```{r include=FALSE}
library(tidyverse)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point()
```

Exercises: Re-create the R code necessary to generate the following graphs:

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() +
  geom_smooth(se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() +
  geom_smooth(se = FALSE, mapping = aes(z = drv))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv)) +
  geom_smooth(se = FALSE, mapping = aes(color = drv))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv)) +
  geom_smooth(se = FALSE)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv), se = FALSE)
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(size = 4, color = "white") +
  geom_point(aes(colour = drv)) 
```

## Statistical Transformation

### Bar Charts

Bar charts seem simple, but they are interesting because they reveal something subtle about plats. The chart shows that more diamonds are available with high-quality cuts than low quality cuts.

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
```

Where does `count` come from? Many graphs, like scatterplots, plot the raw values of your dataset. Other graphs, like bar charts, calculate new values to plot:

  * Bar charts, histograms, and frequency polygons bin your data and then plot bin counts, the number of points that fall in each bin. 
  
  * Smoothers fit a model to your data and then plot predictions from the model. 
  
  * Boxplots compute a robust summary of the distribution and display a specially formatted box. 
  
The algorithm used to calculate new values for a graph is called `stat`, short for statistical transformation. 

You can learn which stat a geom uses by inspecting the default value for the `stat` argument. For example, `?geom_bar()` uses `stat_count()`. `stat_count()` is documented on the same page as `geom_bar()`, and if you scroll down you can find a section called "Computed variables." That tells that it computes two new variables: `count` and `prop`. 

You can generally use geoms and stats interchangeably. For example, you can re-create the previous plot using `stat_count()` instead of `geom_bar()`.

```{r}
ggplot(data = diamonds) +
  stat_count(mapping = aes(x = cut))
```

This works because every geom has a default stat, and every stat has a default geom. This means that you can typically use geoms without worrying about the underlying statistical transformation. There are three reasons you might need to use a stat explicitly:

1. You might want to override the default stat. 
2. You might want to oveeride the deafult mapping from transformed variables to aesthetics. 
3. You might want to draw greater attention to the statistical transformation in your code. 

## Exercises

1. What is the default geom associated with `stat_summary()`? How could you rewrite the previous plot to use that geom function instead of the stat function? 

The default geom associated with `stat_summary()` is `pointrange`. 

```{r}
ggplot(data = diamonds) +
  geom_pointrange(
    mapping = aes(x = cut, y = depth),
    stat = "summary",
    fun.min = min,
    fun.max = max,
    fun = median
  )
```

2. What does `geom_col()` do? How is it different to `geom_bar()`?

The geom_col() function has different default stat than geom_bar(). The default stat of geom_col() is stat_identity(), which leaves the data as is. The geom_col() function expects that the data contains x values and y values which represent the bar height.

The default stat of geom_bar() is stat_count(). The geom_bar() function only expects an x variable. The stat, stat_count(), preprocesses input data by counting the number of observations for each value of x. The y aesthetic uses the values of these counts.

3. Most geoms ans stats come in paris that are almost always used in concert. Read therough the documentation and make a list of all the pairs. What do they have in common?
  
|         geom          |           stat          |
| :-------------------: | :---------------------: |
| geom_bar()	          | stat_count()            |
| geom_bin2d()	        | stat_bin_2d()           |
| geom_boxplot()	      | stat_boxplot()          |
| geom_contour_filled() |	stat_contour_filled()   |
| geom_contour()	      | stat_contour()          |
| geom_count()	        | stat_sum()              |
| geom_density_2d()	    | stat_density_2d()       |
| geom_density()	      | stat_density()          |
| geom_dotplot()	      | stat_bindot()           |
| geom_function()	      | stat_function()         |
| geom_sf()	            | stat_sf()               |
| geom_sf()	            | stat_sf()               |
| geom_smooth()	        | stat_smooth()           |
| geom_violin()	        | stat_ydensity()         |
| geom_hex()	          | stat_bin_hex()          |
| geom_qq_line()	      | stat_qq_line()          |
| geom_qq()	            | stat_qq()               |
| geom_quantile()	      | stat_quantile()         | 



