---
title: "Homework 5"
author: "Yuvraj Jain"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  1,3 (Pg. 227); 2 (Pg. 232); 1,2 (Pg. 235); 

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
library(forcats)
```

# Chapter 12 (Pg. 227)

##  Exercise 1

Explore the distribution of `rincome` (reported income). What makes the default bar chart hard to understand? How could you improve the plot?

```{r}
ggplot(data = gss_cat) +
  geom_bar(mapping = aes(rincome))
```

The problem with the default implementation of `geom_bar()` is hard to understand because it puts the `rincome` variable on the x axis by default, which overlaps text on each other. We can improve this by putting `rincome` on the y axis. Also, we can arrange the income to go in increasing order from the bottom of the graph to the top using factors, which is easier to read.

```{r}
ggplot(data = gss_cat) +
  geom_bar(mapping = aes(y = fct_rev(fct_relevel(rincome, "Not applicable")))) +
  theme_minimal()

```

##  Exercise 3:

Which relig does denom (denomination) apply to? How can you find out with a table? How can you find out with a visualization?

Making the table:

```{r}
gss_cat %>%
  filter(denom != "Not applicable") %>%
  group_by(relig) %>%
  count()
```

From the table we can see over 10,000 rows for Protestant, which tells us that denom primarily applies to the Protestant religion. 

Seeing the same result through a scatterplot.

```{r}
gss_cat %>%
  count(relig, denom) %>%
  ggplot(aes(x = relig, y = denom, size = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90))
```

From the scatterplot, we can see that `denom` refers to Protestant.

# Chapter 12 (Pg. 232)

##  Exercise 2:

For each factor in `gss_cat` identify whether the order of the levels is arbitrary (Nominal) or principled (Ordinal).

There are 6 categorical variables in `gss_cat`:

### 1. `marital`

```{r}
levels(gss_cat$marital)
```

Marital status is somewhat principled where we move from "Never married" to "Separated", "Divorced", "Widowed", and then "Married". But, I would still consider this variable to be mostly arbitrary because there is no clear order to arrange these levels.

### 2. `race`

```{r}
levels(gss_cat$race)
gss_cat %>%
  ggplot(aes(race)) +
  geom_bar() +
  scale_x_discrete(drop = FALSE)
```

The ordering of race is principled in that the categories are ordered by count of observations in the data.

### 3. `rincome`

The levels of rincome are ordered in decreasing order of the income; however the placement of “No answer”, “Don’t know”, and “Refused” before, and “Not applicable” after the income levels is arbitrary

```{r}
levels(gss_cat$rincome)
```

### 4. `relig`

The levels of relig is arbitrary, since there is no way to naturally order the religions. 

### 5. `denom`

The levels of denom is also arbitrary, because, again, we can't order the denominations in any natural order.

### 6. `partyid`

Ignoring “No answer”, “Don’t know”, and “Other party”, the levels of partyid are ordered from “Strong Republican”" to “Strong Democrat”.

```{r}
levels(gss_cat$partyid)
```

# Chapter 12 (Pg. 235)

##  Exercise 1

How have the proportions of people identifying as Democrat, Republican, and Independent changed over time?

```{r}
gss_cat %>%
  mutate(
    partyid =
      fct_collapse(partyid,
        Other = c("No answer", "Don't know", "Other party"),
        Republican = c("Strong republican", "Not str republican"),
        Independent = c("Ind,near rep", "Independent", "Ind,near dem"),
        Democrat = c("Not str democrat", "Strong democrat")
      )
  ) %>%
  count(year, partyid) %>%
  group_by(year) %>%
  mutate(proportion = n / sum(n)) %>%
  ggplot(aes(
    x = year, y = proportion,
    colour = fct_reorder2(partyid, year, proportion)
  )) +
  geom_point() +
  geom_line() +
  labs(colour = "Party ID") +
  theme_minimal() +
  xlab("Year") +
  ylab("Proportion")
```

##  Exercise 2:

How could you collapse `rincome` into a small set of categories?

```{r}
levels(gss_cat$rincome)
```

```{r}
gss_cat %>%
  mutate(
    rincome =
      fct_collapse(
        rincome,
        `Unknown` = c("No answer", "Don't know", "Refused", "Not applicable"),
        `Lt $5000` = c("Lt $1000", "$1000 to 2999", "$3000 to 3999", "$4000 to 4999"),
        `$5000 - $10000` = c("$5000 to 5999", "$6000 to 6999", "$7000 to 7999", "$8000 to 9999"),
        `$10000 - $20000` = c("$10000 - 14999", "$15000 - 19999"),
        `$20000 or more` = c("$25000 or more", "$20000 - 24999")
      )
  ) %>%
  ggplot(aes(x = rincome)) +
  geom_bar() +
  coord_flip()
```
