---
title: "Analysis 2: Connecting Country Level Data"
author: "Yuvraj Jain"
date:  "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

# Instructions

**Overview:** For each question, show your R code that you used to answer each question in the provided chunks. When a written response is required, be sure to answer the entire question in complete sentences outside the code chunks. When figures are required, be sure to follow all requirements to receive full credit. Point values are assigned for every part of this analysis.

**Helpful:** Make sure you knit the document as you go through the assignment. Check all your results in the created HTML file.

**Submission:** Submit via an electronic document on Sakai. Must be submitted as an HTML file generated in RStudio. 

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
library(tidyverse) 
library(rvest)
library(curl) #Make sure you install this package if you don't have it.
```

# Introduction

Many times in data science, your data will be split between many different sources, some of which may be online. In this analysis assignment, we will webscrape country level data from multiple websites, clean the data individually, and merge the data. The website [Worldometers](https://www.worldometers.info/) contains very interesting country level data that when connected may allow us to learn interesting things about the wonderful world in which we exist:

Follow the steps to accomplish specific tasks, and do not modify code on lines with the comment **#DO NOT CHANGE**. Add code in R code chunks wherever you see COMPLETE. After completing each code chunk, change `eval=FALSE` to `eval=TRUE` and **knit** the document.

# Assignment

## Part 1: GDP by Country

Information at [Worldometer GDP](https://www.worldometers.info/gdp/gdp-by-country/) contains GDP data from 2017 published by the world bank. GDP is the monetary value of goods and services produced within a country over a period of time. On this website, GDP is presented in dollars.
 
### Q1 *(3 Points)*

Webscrape the data from https://www.worldometers.info/gdp/gdp-by-country/ into a data frame in R called `GDP`. If done correctly, you should have a new object in R called `GDP` which is a data frame with 190 observations and 8 variables.

```{r,eval=TRUE}
URL.GDP="https://www.worldometers.info/gdp/gdp-by-country/" #DO NOT CHANGE

GDP = read_html(URL.GDP) %>%
  html_table(fill = TRUE) %>%
  .[[1]]

head(GDP) #DO NOT CHANGE
```

### Q2 *(2 Points)*

Remove the first and fourth variables from `GDP` and create a new data frame named `GDP2` based on this change.

```{r,eval=TRUE}
GDP2 = GDP %>%
  select(-c('#', 'GDP (abbrev.)'))

head(GDP2) #DO NOT CHANGE
```

### Q3 *(3 Points)*

Create a new data frame named `GDP3` based off `GDP2` where the variables `GDP (nominal, 2017)`,`GDP  growth`, `Population (2017)`, `GDP  per capita`, and `Share of World GDP` to `GDP`, `Growth`, `Population`, `PerCapita`, and `Share`, respectively. Be careful!! In the original variable names, there are two spaces between "GDP" and "growth" in `GDP growth` and two spaces between "GDP" and "per capita" in GDP  per capita`.

```{r,eval=TRUE}
GDP3 = GDP2 %>%
  rename(
    'GDP' = 'GDP (nominal, 2017)',
    'Growth' = 'GDP  growth',
    'Population' = 'Population (2017)',
    'PerCapita' = 'GDP  per capita',
    'Share' = 'Share of World GDP' 
  )

names(GDP3) #DO NOT CHANGE
```


### Q4 *(3 Points)*

Next, we must clean the data so there are no dollar signs or percent signs in the data using `str_replace()`. The dollar sign is a special character and must be referenced as `\\$`. Create a new data frame named `GDP4` where the dollar signs and percent signs are removed from all necessary variables.

```{r, eval=TRUE}

GDP4 <- GDP3 %>%
  mutate(GDP = str_replace(GDP, pattern = '\\$', replacement = "")) %>%
  mutate(Growth = str_replace(Growth, pattern = "%", replacement = "")) %>%
  mutate(PerCapita = str_replace(PerCapita, pattern = '\\$', replacement = "")) %>%
  mutate(Share = str_replace(Share, pattern = "%", replacement = "")) 

str(GDP4) #DO NOT CHANGE
```

### Q5 *(3 Points)*

Next, create a new data frame named `GDP5` where all commas are removed from potentially numeric variables using `str_replace_all()`.

```{r, eval=TRUE}

GDP5 <- GDP4 %>%
  mutate(GDP = str_replace_all(GDP, pattern = ",", "")) %>%
  mutate(Population = str_replace_all(Population, pattern = ",", "")) %>%
  mutate(PerCapita = str_replace_all(PerCapita, pattern = ",", ""))

str(GDP5) #DO NOT CHANGE
```

### Q6 *(2 Points)*

Create a new data frame called `GDP6` where all the variables except `Country` are changed to numeric variables.

```{r, eval=TRUE}

GDP6 <- GDP5 %>%
  mutate_at(2:6, as.numeric) 

str(GDP6) #DO NOT CHANGE
```


### Q7 *(2 Points)*

Rewrite over the original `GDP` variable with a new variable called `GDP` that is in **trillions of dollars** rather than in actual dollars. Rewrite over the original `Population` variable with a new variable of the same name that is in **millions of people** rather than in actual people. You are scaling the original variables to change the units without changing the variable names. Save your changes in a new data frame called `GDP7`.

```{r, eval=TRUE}

GDP7 <- GDP6 %>%
  mutate(GDP = GDP / 1e12) %>%
  mutate(Population = Population / 1e6) 

str(GDP7)  #DO NOT CHANGE
```

## Part 2: More Relevant Data for Countries

Check out the Wikipedia page (https://en.wikipedia.org/wiki/Education_Index) which contains the education index for all countries from 1990 to 2019.

Check out the Worldometer page (https://www.worldometers.info/co2-emissions/co2-emissions-by-country/) which contains CO2 data for all countries.

### Q1 *(4 Points)*

Webscrape the data from (https://en.wikipedia.org/wiki/Education_Index) into a data frame in R called `EDU` and only keep the 2 variables with each country's name along with its education index from 2019. Also, rename the variable named "2019" to "ED.Index".

```{r,eval=TRUE}
URL.EDU="https://en.wikipedia.org/wiki/Education_Index" #DO NOT CHANGE

EDU = read_html(URL.EDU) %>%
  html_table(fill = TRUE) %>%
  .[[1]] %>%
  select(c('Country', '2019')) %>%
  rename("ED.Index" = '2019')

head(EDU) #DO NOT CHANGE
```


### Q2 *(4 Points)*

Webscrape the data from (https://www.worldometers.info/co2-emissions/co2-emissions-by-country/) into a data frame in R. Keep only the `Country`, and `CO2 Emissions (tons, 2016)` variables. Rename `CO2 Emissions (tons, 2016)` to `CO2` . Also, make the `CO2` variable a numeric variable. Call the final dataset `CO2`.

Each time you run this code chunk, you may have to run the first line of this code every time to get it to work. You may get some error related to "invalid connection" or "open connection".  You can also try knitting the document every time you make any change to this code chunk to see the output in the HTML file.

```{r,eval=TRUE}
URL.CO2=url("https://www.worldometers.info/co2-emissions/co2-emissions-by-country/") #DO NOT CHANGE

CO2 = read_html(URL.CO2) %>%
  html_table(fill = TRUE) %>%
  .[[1]] %>%
  select(c('Country', 'CO2 Emissions (tons, 2016)')) %>%
  rename('CO2' = 'CO2 Emissions (tons, 2016)') %>%
  mutate(CO2 = str_replace_all(CO2, ",", "")) %>%
  mutate_at('CO2', as.numeric)

str(CO2) #DO NOT CHANGE
```

## Part 3: Merging the Datasets and Follow-up 

Now, we are going to merge the datasets for maximum gains. Make sure you carefully read the instructions for each question. Be very careful in this part of the assignment.

### Q1 *(2 Points)*

The dataset named `GDP7` is our primary dataset. Create a new dataset called `GDP7.EDU` that brings the education data from `EDU` into the dataset `GDP7`. Use a left join only.

```{r,eval=TRUE}

GDP7.EDU <- merge(GDP7, EDU, all.x = TRUE)

head(GDP7.EDU) #DO NOT CHANGE
```


### Q2 *(3 Points)*

How many countries in `GDP7.EDU` have missing values for Education Index? Show code that can be used to answer this question and place your answer in the appropriate space below in a complete sentence. Your code should return a number.

Code *(1 Point)*:
```{r,eval=TRUE}

sum(is.na(GDP7.EDU$ED.Index))

```

Answer *(2 Point)*: (Place your answer here in complete sentences.)

24 countries in `GDP.EDU` have missing values for Education Index. 

### Q3 *(3 Points)*

Closely inspect the original datasets and answer the following questions about `GDP7.EDU` in complete sentences. You can use the code chunk below if needed, but this is not required. If you don't reference the appropriate dataset or you are not specific in your answers, you will get 0 points.

```{r,eval=TRUE}

head(GDP7.EDU, 10)

```

Why is there no education index for Russia in the dataset `GDP7.EDU`?

Answer *(1 Point)*: When merging the datasets, in the left dataset Russia is just 'Russia' whereas in the right dataset, it's actually called 'Russian Federation', which is why when we combine these two datasets we see an NA value for Russia. 

Why is there no education index for South Korea in the dataset `GDP7.EDU`?

Answer *(1 Point)*: In the left Dataset, South Korea is named as 'South Korea' but in the right dataset, it's actually called, 'Korea (Republic of)'. South Korea is also known as Republic of Korea, which is when when we merged these through datasets through a left join, 'South Korea' was not able to find any match in the right dataset resulting in an NA value for education index. 

Why is there no education index for Tuvalu in the dataset `GDP7.EDU`?

Answer *(1 Point)*: Tuvalu just does not exist in the EDU dataset. Tuvalu was earlier known as Ellice Islands, and even that name does not exist in the EDU dataset. This is why when we try to merge the two datasets together, we get an NA value of education index for Tuvalu.


### Q4 *(2 Points)*

Suppose we want to see if countries with higher GDP tend to be countries with higher CO2 emissions. Merge the data from `GDP7` with the the data frame named `CO2` to create a new dataset named `GDP7.CO2` that only contains data for countries that are represented in both datasets. 

```{r,eval=TRUE}

GDP7.CO2 <- merge(GDP7, CO2, all.x = TRUE, all.y = TRUE)

head(GDP7.CO2) #DO NOT CHANGE
```

### Q5 *(4 Points)*

In the dataset `CO2`, there are 209 observations. In the dataset `GDP7`, there are 190 observations. Use a filtering join to create a dataset called `CO3` that shows only CO2 data for the countries in `CO2` that are not represented in `GDP7`. Then, answer the follow-up question in complete sentences.

Code *(2 Points)*:
```{r, eval=TRUE}

CO3 <- anti_join(CO2, GDP7, by = "Country")

head(CO3) #DO NOT CHANGE
```

If there are 209 countries in `CO2` and 190 countries in `GDP7`, why is the number of countries in `CO3` not simply $209-190=19$? Answer in complete sentences and fill free to use the code chunk below if needed.

```{r,eval=TRUE}
GDP8 <- anti_join(GDP7, CO2, by = "Country")
CO3
GDP8
```

Answer *(2 Points)*: When we do the anti join of CO2 with GDP7, we see that there are 28 observations which come up instead of 19. This is because not all 190 countries in GDP7 are also present in CO2. In fact, only 181 countries in GDP7 match up with the CO2 dataset. We can confirm this from the code above where we get 9 observations for running the anti join on GDP7 and CO2. When we take into account these 9 observations, our results starts to make more sense because we're actually doing $209-181 = 28$.
